\chapter{Stage History}

\dateHistory{From 9$^{th}$ February to 13$^{th}$ February 2009 (week 1)}
\begin{itemize}
 \item Familiarization with working environment : KDevelop, cmake, svn... .
 \item XML learning.
 \item Configure cmake to compile Qt classes using slot/signals.
 \item Phase 1 : client-server simple communication.
 \item Beginning of phase 2 : user interface.
 \begin{itemize}
  \item The client gets the XML tree and displays it in a treeview using
Model-View programming.
  \item Double-clicking on a tree item displays its XML attributes and their
values in a message box.
 \end{itemize}
\end{itemize}

\dateHistory{From 16$^{th}$ February to 20$^{th}$ February 2009 (week 2)}
\begin{itemize}
 \item "XML attributes" are now called "options".
 \item Create PanAttributes class to make the attributes display more
convinient.
 \begin{itemize}
  \item Attributes names in a label and attributes values in a line edit.
  \item "Read-only" mode.
 \end{itemize}
 \item GUI conception improvements.
 \item Add log window on the main window, displaying messages and errors coming
from the server.
 \item Client can modify an option (on the server).
 \item Right-clicking on a tree item shows a context menu.
 \item Improve XML protocol.
 \item Documentation.
\end{itemize}

\dateHistory{From 23$^{th}$ February to 27$^{th}$ February 2009 (week 3)}
\begin{itemize}
 \item Client can add an option (on the server).
 \item Improve XML protocol (code more readable).
 \item Adapt project to work with a new XML tree format.
 \item Correct some bugs. 
\end{itemize}

\dateHistory{From 2$^{nd}$ March to 5$^{th}$ March 2009 (week 4)}
\begin{itemize}
 \item Correct a bug occuring when the client tries to connect to the server
when it is not running.
 \item Add basic and advanced modes.
 \item Display a tree item properties.
\end{itemize}

\dateHistory{From 9$^{th}$ March to 13$^{th}$ March 2009 (week 5)}
\begin{itemize}
 \item Add concrete and abstract types features.
 \item Integration of my project into \cf as a plugin.
 \item Add concurrent access protection (mutex) to the server.
 \item First tests of opening a case file on the server.
\end{itemize}

\dateHistory{From 16$^{th}$ March to 20$^{th}$ March 2009 (week 6)}
\begin{itemize}
 \item Network frames building process improvements.
 \item Add a dialog on the client to allow user to select a case file to open
(no directory change allowed).
 \item Improvements to slots management on the client.
 \item Improvements to tree management on the server.
 \item Client can launch a simulation on the server from the client (but it
crashes).
 \item Add dynamic options feature : some options are modifiable during the
simulation.
 \item Add ACK/NACK system.
 \item Documentation.
\end{itemize}

\dateHistory{From 23$^{th}$ March to 27$^{th}$ March 2009 (week 7)}
\begin{itemize}
 \item Simplify XML attributes management in XML parser.
 \item Remove some unused class members.
 \item Documentation.
\end{itemize}

\dateHistory{From 30$^{th}$ March to 3$^{rd}$ April 2009 (week 8)}
\begin{itemize}
 \item Client can launch a server instance through an SSH connection (not
fully functional).
 \item Documentation.
\end{itemize}

\dateHistory{From 6$^{th}$ April to 10$^{th}$ April 2009 (week 9)}
\begin{itemize}
 \item Server creates a thread to run the simulation, so new clients can connect
during the simulation.
 \item From the client, user can choose between disconnect from the server or
shutdown the server.
 \item Modify connection dialog GUI : port number only available if in advanced
mode.
 \item Option descriptions are displayed in a tooltip.
 \item Begin to work on a file browser dialog allowing the client to choose a
file to open on the server.
 \item Documentation.
\end{itemize}

\dateHistory{From 14$^{th}$ April to 17$^{th}$ April 2009 (week 10)}
\begin{itemize}
 \item File browser dialog fully functional (with icons) 
 \item Begin to work on my report.
 \item Documentation.
\end{itemize}

\dateHistory{From 20$^{th}$ April to 24$^{th}$ April 2009 (week 11)}
\begin{itemize}
 \item Improve log window behaviour when server sends long messages. 
 \item Test my project (and make it work) on a real network (not the local
loop).
 \item Improve server launch through an SSH connection and correct bugs
revealed by the test on the real network.
 \item Fix a very old bug with the log window.
 \item Improve server checks when the client does not respect the protocol.
 \item Begin to work on my presentation.
 \item Continue my report.
\end{itemize}

\dateHistory{From 27$^{th}$ April to 30$^{th}$ April 2009 (week 12)}
\begin{itemize}
 \item Continue my presentation.
 \item Continue my report.
\end{itemize}

\dateHistory{From 4$^{th}$ May to 8$^{th}$ May 2009 (week 13)}
\begin{itemize}
 \item Continue my presentation.
 \item Continue my report.
\end{itemize}

\dateHistory{From 11$^{th}$ May to 15$^{th}$ May 2009 (week 14)}
\begin{itemize}
 \item Continue my report.
 \item Correct some mistakes revealed while working on my report.
 \item Documentation.
\end{itemize}

\dateHistory{From 18$^{th}$ May to 22$^{th}$ May 2009 (week 15)}
\begin{itemize}
 \item Continue my report.
 \item Adapt graphical components to the option types.
 \item Documentation.
\end{itemize}

\dateHistory{From 25$^{th}$ May to 29$^{th}$ May 2009 (week 16)}
\begin{itemize}
 \item Finish my report.
 \item Correct some last minor problems.
\end{itemize}
